---
title: "main3"
author: "Joshua Bela"
date: "June 15, 2019"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---
Introduction:
What is the nature of this program?

This program is designed to predict real estate prices given a train set with 91 variables.
The code in this file is simplified, as it refers to an abundance of imported custom functions which are defined in a local r script named 'functions, local.r'.





import libraries
```{r}
library(magrittr)# for pipelines
library(mice) #for multivariate imputation via chained equations
library(dplyr) #for select() and union()
library(plyr)# for the mapvalues function
library(randomForest)
library(doParallel)# for parallel core processing
library(caret)# for tune grid searching
```

import data and functions
These will be named the original train and test sets and will not be modified.
```{r}
data_orig = read.csv("./data/train.csv")
data_test_orig = read.csv("./data/test.csv")
source('./functions, local.r')
```

Print out a list of variables in the test data that contain NA values.
```{r}
hasna(data_orig)
```

declare train and test data
These may be modified, unlike the original data frames.
```{r}
data = data_orig
# rename the response variable as 'y' for convenience
data$y = data$SalePrice
data = data %>% select(-SalePrice)
# remove the Id variables as it is not predictive
data = data %>% select(-Id)

data_test = data_test_orig
```

preliminary feature selection and engineering for train set
```{r}
# In each call of split_data() below, the first argument is a factor and the second argument is the numerical value associated with one of the observed factor levels.
# We are creating new numeric variables for each factor level.  These variables are set to 0 if that factor level does not apply to a particular observation.
# These sets of numeric variables will replace the original 2 variables.

# We are doing this because it turns out that...
# only a small subset of the factor levels have predictive numeric information associated with them.

data = split_data(
  data,
  'MasVnrType',
  'MasVnrArea'
)
data = split_data(
  data,
  'BsmtFinType1',
  'BsmtFinSF1'
)
data = split_data(
  data,
  'BsmtFinType2',
  'BsmtFinSF2'
)
data = split_data(
  data,
  'MiscFeature',
  'MiscVal'
)
```
preliminary feature selection and engineering for test set
```{r}
data_test = split_data(
  data_test,
  'MasVnrType',
  'MasVnrArea'
)
data_test = split_data(
  data_test,
  'BsmtFinType1',
  'BsmtFinSF1'
)
data_test = split_data(
  data_test,
  'BsmtFinType2',
  'BsmtFinSF2'
)
data_test = split_data(
  data_test,
  'MiscFeature',
  'MiscVal'
)

```

Imputation for missing data
```{r}
# For each factor, observations of NA are replaced with a new factor level 'na'.
data = data %>% impute_NA_factor()
data_test = data_test %>% impute_NA_factor()

# For each numeric variable, observations of NA are replaced with 0.
data = data %>% impute_0()
data_test = data_test %>% impute_0()
```

Ensure that the levels of each factor of the test contain the levels of the factors in the train set.
```{r}
data_test = match_levels(data, data_test)
```

If there are any variables in the test set which have 'na' factor level observations and there are no corresponding 'na' observations in the test set to prepare for them, then we will revert the 'na' observations in the test set back to NA values.
```{r}
for(i in colnames(data_test)){
  if(sum(data_test[,i] == 'na') > 0 && sum(data[,i] == 'na') == 0){
    print(i)
    data_test[,i] = ifelse(
      data_test[,i] == 'na',
      NA,
      data_test[,i]
    ) %>% as.factor()
  }
}
```

Next, these NA values will be replaced by MICE imputation.
```{r}
cl <- makePSOCKcluster(6)
registerDoParallel(cl)
set.seed(1)

# Deterministic regression imputation via mice
imp <- mice(data_test, method = "polyreg", m = 1, nnet.MaxNWts = Inf, maxit = 20)
# Store data
data_imp <- complete(imp)# Multiple Imputation

stopCluster(cl) # Turn off parallel processing and free up the cores.
registerDoSEQ()

# this takes a long time, so store the resulting data frame as a CSV
write.csv(
  data_imp,
  "./data/data_test_processed.csv",
  row.names = F
)
```

```{r}
trControl <- trainControl(method = "boot",
                          number = 1,
                          savePredictions = 'all',
                          allowParallel = TRUE
                          )
tuneGrid <- expand.grid(
  mtry = seq(25:35)
)

cl <- makePSOCKcluster(6)
registerDoParallel(cl)
set.seed(1)
model <- train(
  y ~ .,
  data = subset(data, select = -MiscFeature.TenC),
  method = 'rf',
  trControl = trControl,
  tuneGrid = tuneGrid
)
stopCluster(cl) # Turn off parallel processing and free up the cores.
registerDoSEQ()

predictions = predict(model, subset(data, select = -y))
"RMSLE =" %>% paste(sqrt(mean((log(predictions) - log(data$y))^2)))
```

Generate predictions for the test set;
write the predictions to a CSV file.
```{r}
predictions_test = predict(model, data_imp)
length(predictions_test)

write.csv(
  data.frame(Id = data_test$Id, SalePrice = predictions_test),
  "./data/submission.csv",
  row.names = F
)
```












